---
title: "chestR workflow: global Cox to local biomarker maps"
author: "Richard Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{chestR workflow: global Cox to local biomarker maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5
)
```

## Overview

**chestR** re-fits a Cox model at each point on a biomarker grid using kernel
weights. Local coefficients show how treatment (and other) effects vary across
biomarker space. The usual workflow is:

1. Fit a global Cox model.
2. Call `chestr()` on biomarker values.
3. Visualise a coefficient with `plot_chestr()` (alias `plot.chestr()`).

This vignette uses a small simulated survival dataset so it builds without
external files.

## Simulate data and fit a global Cox model

```{r simulate}
library(survival)
library(chestR)

set.seed(20250806)
n <- 120
dat <- data.frame(
  trt = rbinom(n, 1, 0.5),
  cov = rnorm(n),
  biom1 = rnorm(n),
  biom2 = rnorm(n)
)

# Mild treatment-effect modification near the origin of biomarker space
lp <- 0.2 * dat$cov +
  log(0.7) * dat$trt *
  exp(-0.5 * (dat$biom1^2 + dat$biom2^2))
dat$time <- rexp(n, rate = exp(lp))
dat$status <- as.integer(dat$time < 4)
dat$time[dat$status == 0L] <- 4

base <- coxph(Surv(time, status) ~ trt + cov, data = dat)
summary(base)$coefficients
```

## Local estimates with `chestr()`

```{r chestr}
biom <- dat[, c("biom1", "biom2")]
cr <- chestr(
  base,
  biom,
  grid.size = 8,
  method = "legacy",
  kern.adj = 2,
  min_events_per_df = 5
)

head(cr[, c("biom1", "biom2", "trt", "ess_events", "events_per_df", "reliable")])
```

Columns of interest include local coefficients (e.g. `trt`), Kish ESS among
events (`ess_events`), events per model df (`events_per_df`), and `reliable`
(whether the point passed the ESS safeguard).

## Plot local treatment effects

```{r plot}
plot_chestr(
  cr,
  trt.param = "trt",
  base = base,
  biom = biom,
  col.scale = "obs",
  reliable_only = TRUE
)
```

By default only reliable grid points are drawn. Set `reliable_only = FALSE` to
include skipped locations (coefficients remain `NA` and are not plotted).

## Optional note on permutation inference

Pointwise local estimates from `chestr()` are descriptive. Formal inference for
treatment-effect heterogeneity (e.g. global tests that the local treatment
surface is flat) typically needs a permutation or bootstrap procedure that
respects the trial randomisation and the kernel bandwidth. Those routines are
outside the core package API illustrated here; see companion analysis scripts
in the Weighted-Regression project for worked examples.

## Session info

```{r session}
sessionInfo()
```
